Echoes of Economic Downturn: Investigating the Persistent Impact of the Great Recession on Birth Rates Among Young Americans

### 2024-02-13 outputs/paper/paper.qmd
Initial Update for the paper

---
title: "Echoes of Economic Downturn: Investigating the Persistent Impact of the Great Recession on Birth Rates Among Young Americans"
author: 
  - Yuanyi (Leo) Liu
  - Emma Teng
thanks: "Code and data are available at: https://github.com/leoyliu/Analyzing-the-Great-Recession-s-Impact-on-Young-Americans-Birth-Rates/tree/main. A replication of various aspects in this paper are available at: https://doi.org/10.48152/ssrp-srs6-t802"
date: today
date-format: long
abstract: "In our study, we delve into the trend of declining US birth rates among young individuals, particularly in light of the Great Recession's aftermath. By examining similar demographic and cohort factors as Kearney et al. (2022), we extend the investigation to include the potential influences of the feminist movement and advancements in technology and media. Our findings align with Kearney et al., reinforcing the notion of changing priorities among younger generations. This replication highlights the complex interplay of societal shifts and economic conditions on fertility decisions, offering a nuanced understanding of contemporary demographic trends."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# install.packages("knitr")
# install.packages("opendatatoronto")
# install.packages("tidyverse")
# install.packages("here")
# install.packages("tidyr")
# install.packages("dplyr")
# install.packages("kableExtra")

# library(knitr)
# library(opendatatoronto)
# library(tidyverse)
# library(here)
# library(tidyr)
# library(dplyr)
# library(kableExtra)
# library(janitor)
```

@@

The paper is structured to facilitate a comprehensive understanding of the study and its implications. Following @sec-intro, @sec-data presents the data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods. @sec-result discusses the results, elaborating on the observed trends and patterns in crime data. @sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader urban issues, and providing suggestions for future research in this area.

@@

## Variables
To better understand the data, the summary_table dataset was developed to offer a more structured and aggregated view of the crime data. By transforming and summarizing the data into a format that displays crime types against years, this dataset simplifies the task of identifying and analyzing trends over time. The dataset includes various crime categories, such as assault, theft, and other types, with each variable representing the annual count of reported cases for each crime type. This longitudinal data provides a detailed view of the crime trends in Toronto over a decade. 
```{r}

```

We proceeded to organize the dataset into two categories: 'Violent' crimes, which included offenses such as 'Assault' and 'Homicide', and 'Non-Violent' crimes, encompassing acts like 'Theft'. It is particularly valuable for understanding the overall landscape of crime in Toronto, focusing on the comparison between violent and non-violent incidents.
```{r}

```

@@

@@

# Results {#sec-result}
In this section, we delve into the core findings of our study, shedding light on the intricate patterns and trends of crime in Toronto from 2014 to 2023. Our analysis dissects both violent and non-violent crime statistics, revealing insightful correlations and variations over the decade.
```{r fig.width=10, fig.height=5.5}

```

```
\newpage

The red line represents violent crimes, and the blue line represents non-violent crimes. The plot shows distinct trends for both categories over the specified time frame.

* **Violent Crimes (Red Line):** There has been an overall increasing trend in the number of violent crime cases over the years. There are some fluctuations, with a noticeable dip occurring in one of the years, but the general direction is upwards, especially towards the end of the period, indicating a rise in violent crimes or an increase in reporting.

* **Non-Violent Crimes (Blue Line):** The trend in non-violent crime cases shows some variability but does not have as clear an increasing pattern as violent crimes. There are years where the rate has increased or decreased, but the last year shows a notable increase, suggesting a possible upward trend at the end of the period.

Overall, the graph indicates that while both violent and non-violent crimes have seen rises and falls, the increase in violent crimes towards the later years is more pronounced. This could reflect changes in societal conditions, law enforcement practices, or reporting mechanisms.

@@

## Variation in Crime Trends {#sec-first-point}

@@

@@

## Weaknesses and Next Steps

@@

>>>>

---
title: "Echoes of Economic Downturn: Investigating the Persistent Impact of the Great Recession on Birth Rates Among Young Americans"
author: 
  - Yuanyi (Leo) Liu
  - Emma Teng
thanks: "Code and data are available at: https://github.com/leoyliu/Analyzing-the-Great-Recession-s-Impact-on-Young-Americans-Birth-Rates/tree/main. A replication of various aspects in this paper are available at: https://doi.org/10.48152/ssrp-srs6-t802"
date: today
date-format: long
abstract: "In this study, we explore the impact of the Great Recession on birth rates among young people in the United States, utilizing a comprehensive analysis of demographic and economic data. Our findings reveal a significant decline in birth rates within this group during and following the economic downturn, highlighting the intricate relationship between economic stability and reproductive decisions. This research contributes to our understanding of how macroeconomic factors can have profound effects on personal life choices, emphasizing the need for supportive policies targeting young individuals during economic crises."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

install.packages("knitr")
install.packages("opendatatoronto")
install.packages("tidyverse")
install.packages("here")
install.packages("tidyr")
install.packages("dplyr")
install.packages("kableExtra")

library(knitr)
library(opendatatoronto)
library(tidyverse)
library(here)
library(tidyr)
library(dplyr)
library(kableExtra)
# library(janitor)
```

@@

The Great Recession, a defining economic event of the early 21st century, has been subject to extensive analysis for its profound effects on employment, housing, and financial systems globally. However, its impact on demographic trends, particularly birth rates among young people in the United States, has not been thoroughly explored. This paper aims to fill this gap by examining how the economic downturn influenced reproductive decisions among this demographic.

Utilizing a blend of quantitative data analysis and demographic research, we investigated birth rate trends among young adults during and following the recessionary period. Our methodology involved analyzing birth rate data in conjunction with economic indicators such as unemployment rates, income levels, and housing market trends to draw correlations and insights.

Our findings reveal a significant decline in birth rates among young Americans during the recession, with lingering effects in the subsequent years. This trend suggests that economic instability has a profound impact on personal life choices, including the decision to have children. The implications of these findings are far-reaching, affecting not only demographic forecasting but also social policy and economic recovery strategies.

The paper is structured to facilitate a comprehensive understanding of the study and its implications. Following @sec-intro, @sec-data presents the data, detailing the data sources, analytical techniques, and the rationale behind the chosen methods. @sec-result discusses the results, elaborating on the observed trends and patterns in brth rate data. @sec-discussion provides an in-depth discussion of these findings, exploring potential factors influencing these trends, drawing connections to broader socio-economic issues, and providing suggestions for future research in this area.

@@

## Variables
To better understand the data, the summary_table dataset was developed to offer a more structured and aggregated view of the crime data. By transforming and summarizing the data into a format that displays crime types against years, this dataset simplifies the task of identifying and analyzing trends over time. The dataset includes various crime categories, such as assault, theft, and other types, with each variable representing the annual count of reported cases for each crime type. This longitudinal data provides a detailed view of the crime trends in Toronto over a decade.

TODO: Add a table here to better understand the data.

ggplot(figure,
        aes(x = year, y = brate_all)) + 
        geom_line(color="red") +
        geom_vline(xintercept = 2007, color = "black", linetype = "dashed") +
        annotate("text", x = 2007, y = 60, label = "Global Financial Crisis of 2007-2008", 
                 color = "black", angle = 90, vjust = -0.5) +
        labs(title = "Birth Rate in America overtime",
            x = "Years",
            y = "Birth rate") +
        theme(panel.background = element_rect(fill = "white"),
              plot.background = element_rect(fill = "white", color = NA),
              plot.title = element_text(color = "black", size = 14, face = "bold"),
              axis.title = element_text(color = "black"))
```
// TODO: Visualization of the data for intuition

@@

// TODO: Add a summary table to show measurements

@@

# Results {#sec-result}
In this section, we delve into the core findings of our study, shedding light on the intricate patterns and trends of crime in Toronto from 2014 to 2023. Our analysis dissects both violent and non-violent crime statistics, revealing insightful correlations and variations over the decade.

```{r}
ggplot(cleaned_data, aes(x = year, y = brate, color = age_group)) +
  geom_line(size=1) +
  geom_vline(xintercept = 2007, color = "black", linetype = "dashed") + 
  annotate("text", x = 2007, y = 60, label = "2007", 
           color = "black", angle = 90, vjust = -0.5) +
  labs(title = "Birth Rate by Age Group Over Time",
       x = "Year",
       y = "Birth Rate") +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white", color = NA),
        plot.title = element_text(color = "black", size = 14, face = "bold"),
        axis.title = element_text(color = "black"))
```
// TODO: Explain the figure

// TODO: Transit to following figure
```{r fig.width=10, fig.height=5.5}
ggplot(merged_data, aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(fill = brate1544_thsnds_ch_pct)) +
  scale_fill_gradient(low = 'darkred', high = '#FFCCCB') +
  labs(title = 'Change in Birth Rate by State Over Time',
       fill = 'Change in Birth Rate') +
  theme_void()
```
// TODO: Explain the figure

@@

## Findings

@@

## Economic Impact Insights

@@

## Societal and Technological Influences

// TODO

## Weaknesses

@@

## Future Research Directions
